---
title: "Chicago"
author: "Zhengkai Fu"
date: "23/05/2020"
output: html_document
---

```{r library, message = FALSE, warning = FALSE}
library(tidyverse)
library(tseries)
library(tsbox) # transform data into time series
library(xts)
library(COVID19) # to get data about covid 19
library(aTSA) # adf.test
library(forecast) #arima model
library(vars) #VAR and Causality
```

This document is to study the possible relationship between COVID-19 and frequency of crime committed in teh city of Chicago.

# COVID19

## Load COVID 19 data for Chicago
```{r covid 19 for chicago region}
covid19_CH <- covid19("USA", level = 3) %>%
  # this cook county contains chicago
  filter(city == "Cook",
         state == "Illinois" ) %>%
  # filter out days when confirmed is zero or one
  # becasue when it was 2 for a very long time
  filter(confirmed > 2)

# brief display
head(covid19_CH)
```

## Overview of the data
### Visualization
```{r overview of covid, message=FALSE}
# plot cumulative cases
# extract for transforming into time series
ts_CH <- covid19_CH %>%
  dplyr::select(date, confirmed) %>%
  ts_xts()

plot.xts(ts_CH,
         main = "Cumulative confirmed cases
         of COVID 19 in Chicago")

# plot daily cases
# first difference
ts_diff_CH <- na.omit(diff(ts_CH))

plot.xts(ts_diff_CH,
         main = "Daily confirmed cases of
         COVID19 in Chicago")
```

# Chicago Crime
## Load crime data
### 2014 - 2020
```{r get chicago data, 2014 - 2020}
chicago <- RSocrata::read.socrata(
  "https://data.cityofchicago.org/resource/ijzp-q8t2.csv?$where=year >= 2014",
  app_token = "hPU78MH7zKApdpUv4OVCInPOQ")

head(chicago)

# add date
chicago <- chicago %>%
  mutate(Date = substr(date, start = 1, stop = 10)) %>%
  mutate(y_month  = substr(date, start = 1, stop = 7)) %>%
  mutate(month = substr(date, start = 6, stop = 7))
```


```{r}
# summary of all crime
chicago_summary <- chicago %>%
  group_by(primary_type) %>%
  summarise(number_of_crime = n()) %>%
  arrange(desc(number_of_crime))

# bar chart of 5 most frequent crime over the years
chicago %>%
  filter(primary_type %in% head(chicago_summary$primary_type, 5)) %>%
  ggplot(aes(x=primary_type, fill=primary_type)) +
  geom_bar(width = 0.5) +
  coord_flip() +
  theme_classic() +
  labs(y='Number of Crimes',x='Offense Description')

# per month
# exclude May
chicago %>%
  dplyr::select(month, primary_type, year) %>%
  filter(primary_type %in% chicago_summary$primary_type[1:5], year == 2020) %>% 
  count(month, primary_type) %>%
  na.omit() %>%
  ggplot(aes(month, n, group = primary_type, color = primary_type)) +
  geom_line() +
  scale_color_brewer(palette = "Set1") +
  ggtitle("Monthly frequency of top 5 crime in chicago in 2020")

# This dataset doesn't have 'day of the week' yet.

# daily
chicago %>%
  dplyr::select(Date, primary_type, year) %>%
  filter(primary_type %in% chicago_summary$primary_type[1:5], year == 2020) %>%
  count(Date, primary_type) %>%
  na.omit() %>%
  ggplot(aes(Date, n, group = primary_type, color = primary_type)) +
  geom_line() +
  facet_wrap(~primary_type) +
  scale_fill_brewer(palette = "Set1", breaks = rev(levels(chicago_summary$primary_type[1:5]))) +
  ggtitle("Frequency of top 5 crime in Chicago in 2020")

# looks life theft is seeing sharp drop

# year to year comparison
chicago %>%
  dplyr::select(y_month, month, primary_type, year) %>%
  filter(primary_type %in% chicago_summary$primary_type[1:5]) %>%
  count(year, month, primary_type) %>%
  na.omit() %>%
  ggplot(aes(x=month, y=n, group = year, color = as.character(year))) +
  geom_line() +
  facet_wrap(~primary_type) +
  ggtitle("year to year comparison of the top 5 crime in Chicago in the past 6 years")
```

## VAR 
### Step1 : Extract cases
```{r extract cases}
# extract top 5 crime
top5crime <- chicago %>%
  filter(primary_type %in% head(chicago_summary$primary_type, 5)) %>%
  group_by(Date, primary_type) %>%
  tally() %>%
  spread(primary_type, n)

# rename columns
colnames(top5crime) <- c('time',
                         "assault",
                         "battery",
                         "criminal",
                         'deceptive',
                         "theft")
top5crime <- na.omit(top5crime)
# create time series
top5crime_xts <- ts_xts(top5crime[,1:2])

for (i in (3:ncol(top5crime))){
  temp_xts <- ts_xts(top5crime[, c(1,i)])
  top5crime_xts <- merge(top5crime_xts, temp_xts)
}

# extract difference, change per day
top5crime_diff <- na.omit(diff(top5crime_xts))
```

### Step 2: Construct combined time series
## COVID 19 RELATED

```{r covid 19 related exploration}
# extract for tranforming into time series data
ts_CH <- covid19_IL %>% 
  dplyr::select(date, confirmed) %>%
  ts_xts()

# plot time series of IL infection
ts_plot(ts_IL)
# conduct ADF Test
adf.test(as.ts(ts_IL))
# not stationary!

# try first log difference
ts_diff_IL <- diff(ts_IL)
ts_plot(ts_diff_IL)
# still clearly not stationary
# need de-trend

# de-trend 
# GAMM model from STA303 A3

# time as integer
covid19_IL_diff$timeInt = as.numeric(covid19_IL_diff$date)
# make a copy to avoid perfect collinearity
covid19_IL_diff$timeIid = covid19_IL_diff$timeInt

# make a copy to avoid perfect collinearity
covid19_IL_diff$timeIid = covid19_IL_diff$timeInt

# GAMM model
# 50 too overfit. 15 looks decent
gamIL <- gamm4::gamm4(confirmed ~  s(timeInt, k=50), random = ~(1|timeIid), 
	data=covid19_IL_diff, family=poisson(link='log'))

lme4::VarCorr(gamIL$mer)
# looks like random intercept is making little difference.
# choose to not have random effect to preserve it for time series analysis

# plot fitted value
toPredict = data.frame(time = seq(covid19_IL_diff$date[1], 
                                          covid19_IL_diff$date[length(covid19_IL_diff$date)],
                                  by = '1 day'))
toPredict$timeInt = as.numeric(toPredict$time)

# obtain forecast
forecast <- data.frame(exp(do.call(cbind, mgcv::predict.gam(gamIL$gam, toPredict, se.fit=TRUE))))

# access residuals
IL_res <- data.frame(covid19_IL_diff$confirmed - forecast$fit)

# transform into time series
IL_res$time = covid19_IL_diff$date
colnames(IL_res)[1] = "residuals"

col_order <- c("time", "residuals")
IL_res <- IL_res[, col_order]

IL_res_ts <- ts_xts(IL_res)

plot.xts(IL_res_ts)
# adf test
adf.test(as.ts(IL_res_ts))
# Stationary process
```

```{r top 5 crime VAR}
# specify common time range
# start from when covid was a thing
# end with crime since it is manually updated
common_time <- seq.Date(start(IL_res_ts), end(top5crime_diff), by = "day")

# combine time series of crime and covid
combined_diff <- merge(top5crime_diff[paste(common_time[1],
                                            common_time[length(common_time)],
                                            sep = "/")],
                       IL_res_ts[paste(common_time[1],
                                            common_time[length(common_time)],
                                            sep = "/")])

```

### Step 3: Plot each crime with covid
```{r plot together}
for (i in 1:(ncol(combined_diff) - 1)){
  plotrix::twoord.plot(common_time,
                       combined_diff[,i],
                       common_time,
                       combined_diff$residuals,
                       type = c("l","l"),
                       xaxt = "n",
                       rylab = "number of daily fluctuation of covid 19 cases",
                       ylab = paste("daily change in", colnames(combined_diff)[i]))
                       
}
```

### Step 5: Construct VAR model
```{r construct var}
optimal_assault <- VARselect(na.omit(combined_diff)[,c(1,6)], type = 'none', lag.max = 10)
optimal_battery <- VARselect(na.omit(combined_diff)[,c(2,6)], type = 'none', lag.max = 10)
optimal_criminal <- VARselect(na.omit(combined_diff)[,c(3,6)], type = 'none', lag.max = 10)
optimal_deceptive <- VARselect(na.omit(combined_diff)[,c(4,6)], type = 'none', lag.max = 10)
optimal_theft <- VARselect(na.omit(combined_diff)[,c(5,6)], type = 'none', lag.max = 10)

VAR_assault <- VAR(y=as.ts(na.omit(combined_diff)[,c(1,6)]), p=optimal_assault$selection[1])
VAR_battery <- VAR(y=as.ts(na.omit(combined_diff)[,c(2,6)]), p=optimal_battery$selection[1])
VAR_criminal <- VAR(y=as.ts(na.omit(combined_diff)[,c(3,6)]),
                    p=optimal_criminal$selection[1])
VAR_deceptive <- VAR(y=as.ts(na.omit(combined_diff)[,c(4,6)]),
                     p=optimal_deceptive$selection[1])
VAR_theft <- VAR(y=as.ts(na.omit(combined_diff)[,c(5,6)]), p=optimal_theft$selection[1])

summary(VAR_assault)
```

### Step 6: Granger Causality test

```{r}
causality(VAR_assault, cause = colnames(combined_diff)[1])
causality(VAR_assault, cause = "residuals")
```
COVID-19 may help to predict assault (p-value = 0.05525)

```{r}
causality(VAR_battery, cause = colnames(combined_diff)[2])
causality(VAR_battery, cause = "residuals")
```
```{r}
causality(VAR_criminal, cause = colnames(combined_diff)[3])
causality(VAR_criminal, cause = "residuals")
```
Very significant causality on criminal damage (p-value = 0.0009405)
instantaneous causality between: residuals and criminal (p-value = 0.01247)

```{r}
causality(VAR_deceptive, cause = colnames(combined_diff)[4])
causality(VAR_deceptive, cause = "residuals")
```

```{r}
causality(VAR_theft, cause = colnames(combined_diff)[5])
causality(VAR_theft, cause = "residuals")
```
COVID-19 may help to predict theft. (p-value = 0.03141)

```{r}
vehicle <- chicago %>%
  filter(Primary.Type == 'MOTOR VEHICLE THEFT')%>%
  group_by(date, Primary.Type) %>%
  tally() %>%
  spread(Primary.Type, n)

colnames(vehicle) <- c('time', 'vehicle')
vehicle_xts <- ts_xts(vehicle[,1:2])
vehicle_diff <- na.omit(diff(vehicle_xts))
```

```{r}
combined_diff2 <- merge(vehicle_diff[paste(common_time[1],
                                            common_time[length(common_time)],
                                            sep = "/")],
                       IL_res_ts[paste(common_time[1],
                                            common_time[length(common_time)],
                                            sep = "/")])
optimal_vehicle <- VARselect(na.omit(combined_diff2)[,c(1,2)], type = 'none', lag.max = 10)
VAR_vehicle <- VAR(y=as.ts(na.omit(combined_diff2)[,c(1,2)]), p=optimal_vehicle$selection[1])

causality(VAR_vehicle, cause = colnames(combined_diff2)[1])
causality(VAR_vehicle, cause = "residuals")
```

### Step 7: Impulse Response Function

Assault, criminal damage and theft have granger causality.

```{r irf}
par(mfrow = c(1,2))
irf_criminal_1 <- irf(VAR_criminal, 
                         impulse = "criminal", 
                         response = "residuals", 
                         n.ahead = 24,
                         ortho = F)
irf_criminal_2 <- irf(VAR_criminal, 
                         impulse = "residuals", 
                         response = "criminal", 
                         n.ahead = 24,
                         ortho = F)
plot(irf_criminal_1)
plot(irf_criminal_2)

# theft
irf_theft_1 <- irf(VAR_theft, 
                         impulse = "theft", 
                         response = "residuals", 
                         n.ahead = 24,
                         ortho = F)
irf_theft_2 <- irf(VAR_theft, 
                         impulse = "residuals", 
                         response = "theft", 
                         n.ahead = 24,
                         ortho = F)
plot(irf_theft_1)
plot(irf_theft_2)

# Calculate impulse response
ir.criminal1 <- irf(VAR_criminal,impulse = "criminal", 
                         response = "residuals", 
                         n.ahead = 24,
                         ortho = F,
                         cumulative = TRUE)

# Plot
plot(ir.criminal1)

ir.criminal2 <- irf(VAR_criminal, 
                         impulse = "residuals", 
                         response = "criminal", 
                         n.ahead = 24,
                         ortho = F,
                         cumulative = T)
plot(ir.criminal2)

ir.theft1 <- irf(VAR_theft, 
                         impulse = "theft", 
                         response = "residuals", 
                         n.ahead = 24,
                         ortho = F,
                         cumulative = T)
plot(ir.theft1)

ir.theft2 <- irf(VAR_theft, 
                         impulse = "residuals", 
                         response = "theft", 
                         n.ahead = 24,
                         ortho = F,
                         cumulative = T)
plot(ir.theft2)
```
### CONCLUSION (Last updated: MAY 30)
Significant contribution on prediction:
Assault (p-value = 0.05525), criminal damage (p-value = 0.0009405), theft (p-value = 0.03141)

Significant simultaneous movement:
Criminal damage (p-value = 0.01247)
  
==================================
