---
title: "LA"
author: "Zhuyi Huang"
output: pdf_document
---
### 2020 ONLY
```{r get LA data, 2020 only}
LA <- read.socrata(
  'http://data.lacity.org/resource/2nrs-mtv8.csv',
  app_token = "hPU78MH7zKApdpUv4OVCInPOQ")

head(LA)

# add date
LA <- LA %>%
  mutate(y_month  = substr(date_occ, start = 1, stop = 7)) %>%
  mutate(MONTH = substr(date_occ, start = 7, stop = 7))
```

```{r}
# summary of all crime
LA_summary <- LA %>%
  group_by(crm_cd_desc) %>%
  summarise(number_of_crime = n()) %>%
  arrange(desc(number_of_crime))

# bar chart of 5 most frequent crime over the years
LA %>%
  filter(crm_cd_desc %in% head(LA_summary$crm_cd_desc, 5)) %>%
  ggplot(aes(x=crm_cd_desc, fill=crm_cd_desc)) +
  geom_bar(width = 0.5) +
  coord_flip() +
  theme_classic() +
  labs(y='Number of Crimes',x='Offense Description')

# per month
# exclude May
LA %>%
  dplyr::select(MONTH, crm_cd_desc) %>%
  filter(crm_cd_desc %in% LA_summary$crm_cd_desc[1:5], MONTH != 05) %>% 
  count(MONTH, crm_cd_desc) %>%
  ggplot(aes(MONTH, n, group = crm_cd_desc, color = crm_cd_desc)) +
  geom_line() +
  scale_color_brewer(palette = "Set1") +
  ggtitle("Monthly frequency of top 5 crime in LA in 2020") +
  theme(plot.title = element_text(hjust = 0.5))

# This data set doesn't have 'day of the week' yet.
```

## VAR 
### Step1 : Extract cases
```{r extract cases}
# extract top 5 crime
top5crime <- LA %>%
  filter(crm_cd_desc %in% head(LA_summary$crm_cd_desc, 5)) %>%
  group_by(date_occ, crm_cd_desc) %>%
  tally() %>%
  spread(crm_cd_desc, n)

# rename columns
colnames(top5crime) <- c('time',
                         "battery",
                         "burglary",
                         'burglary from vehicle',
                         "vandalism",
                         'vehicle')
# create date
top5crime$time <- as.Date(top5crime$time,
                          format = "%Y-%m-%d")
top5crime <- na.omit(top5crime)
# create time series
top5crime_xts <- ts_xts(top5crime[,1:2])

for (i in (3:ncol(top5crime))){
  temp_xts <- ts_xts(top5crime[, c(1,i)])
  top5crime_xts <- merge(top5crime_xts, temp_xts)
}

# extract difference, change per day
top5crime_diff <- na.omit(diff(top5crime_xts))
```

### Step 2: Construct combined time series
## COVID 19 RELATED
```{r covid 19 extract Pennsylvania data}
# extract Pennsylvania data from US data. 
covid19_CA <- covid19("USA", level = 2) %>%
  filter(administrative_area_level_2 == "California") %>%
  # filter out days when confirmed is zero
  filter(confirmed > 0)

# calculate the difference per day
covid19_CA_diff <- data.frame(diff(covid19_CA$confirmed))
colnames(covid19_CA_diff)[1] = "confirmed"
covid19_CA_diff$date = covid19_CA$date[2:length(covid19_CA$date)]

head(covid19_CA)
```

```{r covid 19 related exploration}
# extract for tranforming into time series data
ts_CA <- covid19_CA %>% 
  dplyr::select(date, confirmed) %>%
  ts_xts()

# plot time series of California infection
ts_plot(ts_CA)
# conduct ADF Test
adf.test(as.ts(ts_CA))

# try first log difference
ts_diff_CA <- diff(ts_CA)
ts_plot(ts_diff_CA)
# still clearly not stationary
# need de-trend

# de-trend 
# GAMM model from STA303 A3

# time as integer
covid19_CA_diff$timeInt = as.numeric(covid19_CA_diff$date)
# make a copy to avoid perfect collinearity
covid19_CA_diff$timeIid = covid19_CA_diff$timeInt

# make a copy to avoid perfect collinearity
covid19_CA_diff$timeIid = covid19_CA_diff$timeInt

# GAMM model
# 50 too overfit. 15 looks decent
gamCA <- gamm4::gamm4(confirmed ~  s(timeInt, k=50), random = ~(1|timeIid), 
	data=covid19_CA_diff, family=poisson(link='log'))

lme4::VarCorr(gamCA$mer)
# looks like random intercept is making little difference.
# choose to not have random effect to preserve it for time series analysis

# plot fitted value
toPredict = data.frame(time = seq(covid19_CA_diff$date[1], 
                                          covid19_CA_diff$date[length(covid19_CA_diff$date)],
                                  by = '1 day'))
toPredict$timeInt = as.numeric(toPredict$time)

# obtain forecast
forecast <- data.frame(exp(do.call(cbind, mgcv::predict.gam(gamCA$gam, toPredict, se.fit=TRUE))))

# access residuals
CA_res <- data.frame(covid19_CA_diff$confirmed - forecast$fit)

# transform into time series
CA_res$time = covid19_CA_diff$date
colnames(CA_res)[1] = "residuals"

col_order <- c("time", "residuals")
CA_res <- CA_res[, col_order]

CA_res_ts <- ts_xts(CA_res)

plot.xts(CA_res_ts)
# adf test
adf.test(as.ts(CA_res_ts))
```

```{r top 5 crime VAR}
# specify common time range
# start from when covid was a thing
# end with crime since it is manually updated
common_time <- seq.Date(start(CA_res_ts), end(top5crime_diff), by = "day")

# combine time series of crime and covid
combined_diff <- merge(top5crime_diff[paste(common_time[1],
                                            common_time[length(common_time)],
                                            sep = "/")],
                       CA_res_ts[paste(common_time[1],
                                            common_time[length(common_time)],
                                            sep = "/")])

```

### Step 3: Plot each crime with covid
```{r plot together}
for (i in 1:(ncol(combined_diff) - 1)){
  plotrix::twoord.plot(common_time,
                       combined_diff[,i],
                       common_time,
                       combined_diff$residuals,
                       type = c("l","l"),
                       xaxt = "n",
                       rylab = "number of daily fluctuation of covid 19 cases",
                       ylab = paste("daily change in", colnames(combined_diff)[i]))
                       
}
```

### Step 5: Construct VAR model
```{r construct var}
optimal_battery <- VARselect(na.omit(combined_diff)[,c(1,6)], type = 'none', lag.max = 10)
optimal_burglary <- VARselect(na.omit(combined_diff)[,c(2,6)], type = 'none', lag.max = 10)
optimal_burglary_vehicle <- VARselect(na.omit(combined_diff)[,c(3,6)], type = 'none', lag.max = 10)
optimal_vandalism <- VARselect(na.omit(combined_diff)[,c(4,6)], type = 'none', lag.max = 10)
optimal_vehicle <- VARselect(na.omit(combined_diff)[,c(5,6)], type = 'none', lag.max = 10)

VAR_battery <- VAR(y=as.ts(na.omit(combined_diff)[,c(1,6)]), p=optimal_battery$selection[1])
VAR_burglary <- VAR(y=as.ts(na.omit(combined_diff)[,c(2,6)]),
                    p=optimal_burglary$selection[1])
VAR_burglary_vehicle <- VAR(y=as.ts(na.omit(combined_diff)[,c(3,6)]),
                            p=optimal_burglary_vehicle$selection[1])
VAR_vandalism <- VAR(y=as.ts(na.omit(combined_diff)[,c(4,6)]),
                     p=optimal_vandalism$selection[1])
VAR_vehicle <- VAR(y=as.ts(na.omit(combined_diff)[,c(5,6)]), p=optimal_vehicle$selection[1])
```

### Step 6: Granger Causality test

```{r granger offense}
causality(VAR_battery, cause = colnames(combined_diff)[1])
causality(VAR_battery, cause = "residuals")
```

```{r}
causality(VAR_burglary, cause = colnames(combined_diff)[2])
causality(VAR_burglary, cause = "residuals")
```

```{r}
causality(VAR_burglary_vehicle, cause = colnames(combined_diff)[3])
causality(VAR_burglary_vehicle, cause = "residuals")
```
COVID-19 may help to predict burglary from vehicle (p-value = 0.04295).

```{r}
causality(VAR_vandalism, cause = colnames(combined_diff)[4])
causality(VAR_vandalism, cause = "residuals")
```
COVID-19 may help to predict vandalism (p-value = 0.02398)

```{r}
causality(VAR_vehicle, cause = colnames(combined_diff)[5])
causality(VAR_vehicle, cause = "residuals")
```
### Step 7: Impulse Response Function

burglary from vehicle and vandalism have granger causality.

```{r irf}
par(mfrow = c(1,2))
irf_burg_v_1 <- irf(VAR_burglary_vehicle, 
                         impulse = "burglary.from.vehicle", 
                         response = "residuals", 
                         n.ahead = 24)
irf_burg_v_2 <- irf(VAR_burglary_vehicle, 
                         impulse = "residuals", 
                         response = "burglary.from.vehicle", 
                         n.ahead = 24)
plot(irf_burg_v_1)
plot(irf_burg_v_2)

irf_vandalism_1 <- irf(VAR_vandalism, 
                         impulse = "vandalism", 
                         response = "residuals", 
                         n.ahead = 24)
irf_vandalism_2 <- irf(VAR_vandalism, 
                         impulse = "residuals", 
                         response = "vandalism", 
                         n.ahead = 24)
plot(irf_vandalism_1)
plot(irf_vandalism_2)
```
### CONCLUSION (Last update: MAY 30)
Significant contribution on prediction:
 burglary from vehicle (p-value = 0.04295), vandalism (p-value = 0.02398)

Significant simultaneous movement:
	Not found

  
==================================