---
title: "Providence"
author: "Zhengkai Fu"
date: "21/05/2020"
output: pdf_document
---

```{r library, message = FALSE, warning = FALSE}
library(RSocrata)
library(tidyverse)
library(tsbox) # transform data into time series
library(xts)
library(COVID19) # to get data about covid 19
library(aTSA) # adf.test
library(forecast) #arima model
library(vars) #VAR and Causality
```
# COVID 19
## Load COVID 19 data for Massachusetts
```{r covid 19 for MA}
# extract MA data from US data. level 3 is by cities but cannot find Boston.
covid19_MA <- covid19("USA", level = 2) %>%
  filter(state == "Massachusetts") %>%
  # filter out days when confirmed is zero or one
  # becasue when it was 1 for a very long time
  filter(confirmed > 1)

# brief display
head(covid19_MA)
```
### Choice explained

#### Scope of data

The covid 19 data related to the whole state of Massachusetts are chosen, because the auther believes that suburban area and capital city are closely related in the context of disease and crime. It is well known that in America most of the residential area, suburban or rual area, is separately identifiable from commercial zone, the cities. While the crime might have happened in the city of Boston, the suspects or victims might live outside of Boston.

#### Confirmed cases instead of Death count

Although the number of confirmed cases can largely be influenced by the testing policy of the local government, the psychological effects of seeing confirmed cases alone might be enough to have some psychological relationship with committing crimes, which is what the author would like to explore.

## Overview of the data
### Visualization
```{r overview of covid, message=FALSE}
# plot cumulative cases
# extract for tranforming into time series data
ts_MA <- covid19_MA %>% 
  dplyr::select(date, confirmed) %>%
  ts_xts()

plot.xts(ts_MA,
         main = "Cumulative confirmed cases 
         of COVID19 in Boston")

# plot daily cases
# first difference
ts_diff_MA <- diff(ts_MA)
plot.xts(ts_diff_MA,
         main = "Daily confirmed cases of
         COVID19 in Boston")
```

As shown above, cumulative cases and daily cases have been plotted above.

### Model the infection
#### Construct the model
```{r GAMM model for covid}
# construct GAMM model from prof. Brown's work

# construct data frame of difference, not time series
covid19_MA_diff <- data.frame(diff(covid19_MA$confirmed))
colnames(covid19_MA_diff)[1] = "confirmed"
covid19_MA_diff$date = covid19_MA$date[2:length(covid19_MA$date)]

# time as integer
covid19_MA_diff$timeInt = as.numeric(covid19_MA_diff$date)
# make a copy to avoid perfect collinearity for mixed effect
covid19_MA_diff$timeIid = covid19_MA_diff$timeInt

# GAMM model
gamMA <- gamm4::gamm4(confirmed ~ s(timeInt, k = 70), 
                      random = ~(1|timeIid),
                      data = covid19_MA_diff,
                      family = poisson(link = 'log'))
# currently 70 is the max due to length of data
```

In order to study covid19's impact on Boston, its trend needs to be modeled first to have a better understanding of the situation. A Generalized Additive Mixed Model is used here, which is a direct copy from prof. Patrick Brown's work in STA303 Assignment 3.

#### Visuzalization of the model
```{r plot GAMM}
# plot fitted value
toPredict = data.frame(time = seq(covid19_MA_diff$date[1], 
                                          covid19_MA_diff$date[length(covid19_MA_diff$date)],
                                  by = '1 day'))
toPredict$timeInt = as.numeric(toPredict$time)

# plot

matplot(toPredict$time, 
        exp(do.call(cbind, mgcv::predict.gam(gamMA$gam, toPredict, se.fit=TRUE)) %*% 
              Pmisc::ciMat()), 
        col='red', lty=c(1,2,2), type='l', xaxt='n', xlab='', ylab='Daliy Confirmed cases', 
        ylim = c(0.5, 3500), xlim = as.Date(c(covid19_MA$date[1], covid19_MA$date[length(covid19_MA$date)])))
title("Daily confirmed cases of COVID 19 in Boston")

matpoints(toPredict$time, covid19_MA_diff$confirmed, 
          col = 'black',
          type = 'l')
```
The overall trend of the covid 19 infection in Boston is successfully modeled and displayed above, including a 95% confidence interval. However, those are not very important. Now let's take a look at crime situation in Boston.

#### Obtain the residuals
```{r covid 19 residual}
# obtain forecast
forecast_covid <- data.frame(exp(do.call(cbind, mgcv::predict.gam(gamMA$gam, toPredict, se.fit=TRUE))))
                        
                        
# access residuals
MA_res <- data.frame(covid19_MA_diff$confirmed - forecast_covid$fit)

# transform into time series
MA_res$time = covid19_MA_diff$date
colnames(MA_res)[1] = "residuals"

col_order <- c("time", "residuals")
MA_res <- MA_res[, col_order]

MA_res_ts <- ts_xts(MA_res)

plot.xts(MA_res_ts,
         main = "Daily fluctuation of COVID 19 confirmd cases
         in MA outside its overall trend")
# adf test
adf.test(as.ts(MA_res_ts))

# lag residuals by 4 days due to testing time
MA_res_lag <- stats::lag(MA_res_ts, k = -4)
```

The residuals obtained here represents the fluctuation of daily confirmed cases of COVID 19 in Massachusetts, in order to obtain a stationary time series for future analysis. This residuals can otherwise be interpreted as "fluctuation outside the new norm".
# Providence Crime
## Load crime data
```{r get Providence data}
providence <- read.socrata(
  "https://data.providenceri.gov/resource/rz3y-pz8v.csv",
  app_token = "hPU78MH7zKApdpUv4OVCInPOQ")

head(providence)

# add date
providence <- providence %>%
  mutate(date = substr(reported_date, start = 1, stop = 10)) %>%
  mutate(y_month  = substr(reported_date, start = 1, stop = 7))

# extract most vague name of crime
providence <- providence %>%
  separate(offense_desc, c("crime", "detail_crime"), sep = ",")

```

## Overview of crime
```{r summary}
# summary of all crime
providence_summary <- providence %>%
  group_by(crime) %>%
  summarise(number_of_crime = n()) %>%
  arrange(desc(number_of_crime))

providence_summary

providence %>%
  filter(crime %in% head(providence_summary$crime, 5)) %>%
  ggplot(aes(x=crime)) +
  geom_bar()

# top 5 crime over time
# monthly
# not too many months anyways
providence %>%
  dplyr::select(y_month, crime) %>%
  filter(crime %in% head(providence_summary$crime, 5)) %>% 
  count(y_month, crime) %>%
  ggplot(aes(y_month, n, group = crime, color = crime)) +
  geom_line() + 
  scale_fill_brewer(palette = "Set1", breaks = rev(levels(providence_summary$OFFENSE_DESCRIPTION[1:5]))) +
  ggtitle("Frequency of top 5 crime in Providence")

# daily
providence %>%
  dplyr::select(date, crime) %>%
  filter(crime %in% head(providence_summary$crime, 5)) %>% 
  count(date, crime) %>%
  ggplot(aes(date, n, group = crime, color = crime)) +
  geom_line() + 
  facet_wrap(~crime) +
  scale_fill_brewer(palette = "Set1", breaks = rev(levels(providence_summary$OFFENSE_DESCRIPTION[1:5]))) +
  ggtitle("Frequency of top 5 crime in Providence")
```


## VAR for top 5 crime
### Step 1: Extract cases
```{r extract cases}
# extract top 5 cases
top5crime <- providence %>%
  filter(crime %in% head(providence_summary$crime,5)) %>%
  group_by(date, crime) %>%
  tally() %>%
  spread(crime, n)

# replace NA with 0
top5crime[is.na(top5crime)] = 0

# rename columns
colnames(top5crime) <- c("time",
                         "assault",
                         "larceny",
                         "larceny_from_vehicle",
                         "missing",
                         "vandalism")

# create date
top5crime$time <- as.Date(top5crime$time,
                          format = "%Y-%m-%d")

# create time series
top5crime_xts <- ts_xts(top5crime[,1:2])

for (i in (3:ncol(top5crime))){
  temp_xts <- ts_xts(top5crime[, c(1,i)])
  top5crime_xts <- merge(top5crime_xts, temp_xts)
}

# extract difference, change per day
top5crime_diff <- na.omit(diff(top5crime_xts))
```

### Step 2: Construct combined time series
```{r top 5 crime VAR}
# specify common time range
# start from when covid was a thing
# end with crime since it is manually updated
common_time <- seq.Date(start(MA_res_ts), end(top5crime_diff), by = "day")

# combine time series of crime and covid
combined_diff <- merge(top5crime_diff[paste(common_time[1],
                                            common_time[length(common_time)],
                                            sep = "/")],
                       MA_res_ts[paste(common_time[1],
                                            common_time[length(common_time)],
                                            sep = "/")])

```

### Step 3: Plot each crime with covid
```{r plot together}
for (i in 1:(ncol(combined_diff) - 1)){
  plotrix::twoord.plot(common_time,
                       combined_diff[,i],
                       common_time,
                       combined_diff$residuals,
                       type = c("l","l"),
                       xaxt = "n",
                       rylab = "number of daily fluctuation of covid 19 cases",
                       ylab = paste("daily change in", colnames(combined_diff)[i]))
                       
}
```

Many look very promising

### Step 5: Construct VAR model
```{r construct var}
# variable selection based on AIC
VARselect(combined_diff[,c(1,6)], type = 'none', lag.max = 10)
VARselect(combined_diff[,c(2,6)], type = 'none', lag.max = 10)
VARselect(combined_diff[,c(3,6)], type = 'none', lag.max = 10)
VARselect(combined_diff[,c(4,6)], type = 'none', lag.max = 10)
VARselect(combined_diff[,c(5,6)], type = 'none', lag.max = 10)
```
