---
title: "Providence"
author: "Zhengkai Fu"
date: "21/05/2020"
output: pdf_document
---

```{r library, message = FALSE, warning = FALSE}
library(RSocrata)
library(tidyverse)
library(tsbox) # transform data into time series
library(xts)
library(COVID19) # to get data about covid 19
library(aTSA) # adf.test
library(forecast) #ariRI model
library(vars) #VAR and Causality
```
# COVID 19
## Load COVID 19 data for RIssachusetts
```{r covid 19 for RI}
# extract RI data from US data. level 3 is by cities but cannot find Boston.
covid19_RI <- covid19("USA", level = 2) %>%
  filter(state == "Rhode Island") %>%
  # filter out days when confirmed is zero or one
  # becasue when it was 1 for a very long time
  filter(confirmed > 1)

# brief display
head(covid19_RI)
```
### Choice explained

#### Scope of data

The covid 19 data related to the whole state of RIssachusetts are chosen, because the auther believes that suburban area and capital city are closely related in the context of disease and crime. It is well known that in America most of the residential area, suburban or rual area, is separately identifiable from commercial zone, the cities. While the crime might have happened in the city of Boston, the suspects or victims might live outside of Boston.

#### Confirmed cases instead of Death count

Although the number of confirmed cases can largely be influenced by the testing policy of the local government, the psychological effects of seeing confirmed cases alone might be enough to have some psychological relationship with committing crimes, which is what the author would like to explore.

## Overview of the data
### Visualization
```{r overview of covid, message=FALSE}
# plot cumulative cases
# extract for transforming into time series data
ts_RI <- covid19_RI %>% 
  dplyr::select(date, confirmed) %>%
  ts_xts()

plot.xts(ts_RI,
         main = "Cumulative confirmed cases 
         of COVID19 in Rhode Island")

# plot daily cases
# first difference
ts_diff_RI <- diff(ts_RI)

# since the end goal is to get residuals
# can add 2 to all values so that there is no negative value
# while having residuals unchanged
adj_diff_RI <- na.omit(ts_diff_RI[,1] + 2)

plot.xts(ts_diff_RI,
         main = "Daily confirmed cases of
         COVID19 in Rhode Island")
```

As shown above, cumulative cases and daily cases have been plotted above.

### Model the infection
#### Construct the model
```{r GAMM model for covid}
# construct GAMM model from prof. Brown's work

# construct data frame of difference, not time series
covid19_RI_diff <- data.frame(diff(covid19_RI$confirmed) + 2)
  
colnames(covid19_RI_diff)[1] = "confirmed"
covid19_RI_diff$date = covid19_RI$date[2:length(covid19_RI$date)]

# time as integer
covid19_RI_diff$timeInt = as.numeric(covid19_RI_diff$date)
# RIke a copy to avoid perfect collinearity for mixed effect
covid19_RI_diff$timeIid = covid19_RI_diff$timeInt

# GAMM model
gamRI <- gamm4::gamm4(confirmed ~ s(timeInt, k = 70), 
                      random = ~(1|timeIid),
                      data = covid19_RI_diff,
                      family = poisson(link = 'log'))
# currently 70 is the max due to length of data
```

In order to study covid19's impact on Boston, its trend needs to be modeled first to have a better understanding of the situation. A Generalized Additive Mixed Model is used here, which is a direct copy from prof. Patrick Brown's work in STA303 Assignment 3.

#### Visuzalization of the model
```{r plot GAMM}
# plot fitted value
toPredict = data.frame(time = seq(covid19_RI_diff$date[1], 
                                          covid19_RI_diff$date[length(covid19_RI_diff$date)],
                                  by = '1 day'))
toPredict$timeInt = as.numeric(toPredict$time)

# plot

matplot(toPredict$time, 
        exp(do.call(cbind, mgcv::predict.gam(gamRI$gam, toPredict, se.fit=TRUE)) %*% 
              Pmisc::ciMat()), 
        col='red', lty=c(1,2,2), type='l', xaxt='n', xlab='', ylab='Daliy Confirmed cases', 
        ylim = c(0.5, 800), xlim = as.Date(c(covid19_RI$date[1], covid19_RI$date[length(covid19_RI$date)])))
title("Daily confirmed cases of COVID 19 in Boston")

matpoints(toPredict$time, covid19_RI_diff$confirmed, 
          col = 'black',
          type = 'l')
```
The overall trend of the covid 19 infection in Boston is successfully modeled and displayed above, including a 95% confidence interval. However, those are not very important. Now let's take a look at crime situation in Boston.

#### Obtain the residuals
```{r covid 19 residual}
# obtain forecast
forecast_covid <- data.frame(exp(do.call(cbind, mgcv::predict.gam(gamRI$gam, toPredict, se.fit=TRUE))))
                        
                        
# access residuals
RI_res <- data.frame(covid19_RI_diff$confirmed - forecast_covid$fit)

# transform into time series
RI_res$time = covid19_RI_diff$date
colnames(RI_res)[1] = "residuals"

col_order <- c("time", "residuals")
RI_res <- RI_res[, col_order]

RI_res_ts <- ts_xts(RI_res)

plot.xts(RI_res_ts,
         main = "Daily fluctuation of COVID 19 confirmd cases
         in RI outside its overall trend")
# adf test
adf.test(as.ts(RI_res_ts))
```

The residuals obtained here represents the fluctuation of daily confirmed cases of COVID 19 in providence, in order to obtain a stationary time series for future analysis. This residuals can otherwise be interpreted as "fluctuation outside the new norm".
# Providence Crime
## Load crime data
```{r get Providence data}
providence <- read.socrata(
  "https://data.providenceri.gov/resource/rz3y-pz8v.csv",
  app_token = "hPU78MH7zKApdpUv4OVCInPOQ")

head(providence)

# add date
providence <- providence %>%
  mutate(date = substr(reported_date, start = 1, stop = 10)) %>%
  mutate(y_month  = substr(reported_date, start = 1, stop = 7))

# extract most vague name of crime
providence <- providence %>%
  separate(offense_desc, c("crime", "detail_crime"), sep = ",")

```

## Overview of crime
```{r summary}
# summary of all crime
providence_summary <- providence %>%
  group_by(crime) %>%
  summarise(number_of_crime = n()) %>%
  arrange(desc(number_of_crime))

providence_summary

providence %>%
  filter(crime %in% head(providence_summary$crime, 5)) %>%
  ggplot(aes(x=crime)) +
  geom_bar()

# top 5 crime over time
# monthly
# not too many months anyways
providence %>%
  dplyr::select(y_month, crime) %>%
  filter(crime %in% head(providence_summary$crime, 5)) %>% 
  count(y_month, crime) %>%
  ggplot(aes(y_month, n, group = crime, color = crime)) +
  geom_line() + 
  scale_fill_brewer(palette = "Set1", breaks = rev(levels(providence_summary$crime[1:5]))) +
  ggtitle("Frequency of top 5 crime in Providence")

# daily
providence %>%
  dplyr::select(date, crime) %>%
  filter(crime %in% head(providence_summary$crime, 5)) %>% 
  count(date, crime) %>%
  ggplot(aes(date, n, group = crime, color = crime)) +
  geom_line() + 
  facet_wrap(~crime) +
  scale_fill_brewer(palette = "Set1", breaks = rev(levels(providence_summary$crime[1:5]))) +
  ggtitle("Frequency of top 5 crime in Providence")
```


## VAR for top 5 crime
### Step 1: Extract cases
```{r extract cases}
# extract top 5 cases
top5crime <- providence %>%
  filter(crime %in% head(providence_summary$crime,5)) %>%
  group_by(date, crime) %>%
  tally() %>%
  spread(crime, n)

# replace NA with 0
top5crime[is.na(top5crime)] = 0

# rename columns
colnames(top5crime) <- c("time",
                         "assault",
                         "larceny",
                         "larceny_from_vehicle",
                         "missing",
                         "vandalism")

# create date
top5crime$time <- as.Date(top5crime$time,
                          format = "%Y-%m-%d")

# create time series
top5crime_xts <- ts_xts(top5crime[,1:2])

for (i in (3:ncol(top5crime))){
  temp_xts <- ts_xts(top5crime[, c(1,i)])
  top5crime_xts <- merge(top5crime_xts, temp_xts)
}

# extract difference, change per day
top5crime_diff <- na.omit(diff(top5crime_xts))
```

### Step 2: Construct combined time series
```{r top 5 crime VAR}
# specify common time range
# start from when covid was a thing
# end with 1 day before today's date
common_time <- seq.Date(start(RI_res_ts), Sys.Date() - 1 , by = "day")

# combine time series of crime and covid
combined_diff <- merge(top5crime_diff[paste(common_time[1],
                                            common_time[length(common_time)],
                                            sep = "/")],
                       RI_res_ts[paste(common_time[1],
                                            common_time[length(common_time)],
                                            sep = "/")])

```

### Step 3: Plot each crime with covid
```{r plot together}
for (i in 1:(ncol(combined_diff) - 1)){
  plotrix::twoord.plot(common_time,
                       combined_diff[,i],
                       common_time,
                       combined_diff$residuals,
                       type = c("l","l"),
                       xaxt = "n",
                       rylab = "number of daily fluctuation of covid 19 cases",
                       ylab = paste("daily change in", colnames(combined_diff)[i]))
                       
}
```

Many look very promising

### Step 5: Construct VAR model
```{r construct var, warning = FALSE}
# variable selection based on AIC
optimal_assault <- VARselect(combined_diff[,c(1,6)], type = 'none', lag.max = 10)
optimal_larceny <- VARselect(combined_diff[,c(2,6)], type = 'none', lag.max = 10)
optimal_vehicle <- VARselect(combined_diff[,c(3,6)], type = 'none', lag.max = 10)
optimal_missing <- VARselect(combined_diff[,c(4,6)], type = 'none', lag.max = 10)
optimal_vandalism <- VARselect(combined_diff[,c(5,6)], type = 'none', lag.max = 10)

# use AIC as selection criteria
VAR_assault <- VAR(y=as.ts(combined_diff[,c(1,6)]), p=optimal_assault$selection[1])
VAR_larceny <- VAR(y=as.ts(combined_diff[,c(2,6)]), p=optimal_larceny$selection[1])
VAR_vehicle <- VAR(y=as.ts(combined_diff[,c(3,6)]), p=optimal_vehicle$selection[1])
VAR_missing <- VAR(y=as.ts(combined_diff[,c(4,6)]), p=optimal_missing$selection[1])
VAR_vandalism <- VAR(y=as.ts(combined_diff[,c(5,6)]), p=optimal_vandalism$selection[1])
```

### Step 6: Granger Causality test
#### ASSAULT
```{r granger assault, warning=FALSE}
causality(VAR_assault, cause = colnames(combined_diff)[1])
causality(VAR_assault, cause = "residuals")
```
Nothing significant

#### LARCENY
```{r granger larceny, warning=FALSE}
causality(VAR_larceny, cause = colnames(combined_diff)[2])
causality(VAR_larceny, cause = "residuals")
```
Nothing significant

#### LARCENY FROM VEHICLE
```{r granger vehicle, warning=FALSE}
causality(VAR_vehicle, cause = colnames(combined_diff)[3])
causality(VAR_vehicle, cause = "residuals")
```

Larceny from vehicle significant to covid.
covid significant to larceny from vehicle, weak.

#### MISSING
```{r granger missing, warning=FALSE}
causality(VAR_missing, cause = colnames(combined_diff)[4])
causality(VAR_missing, cause = "residuals")
```
Nothing significant

#### VANDALISM
```{r granger vandalism, warning=FALSE}
causality(VAR_vandalism, cause = colnames(combined_diff)[5])
causality(VAR_vandalism, cause = "residuals")
```
Not significant.

### Step 7: Impulse Response Function

Only larceny from vechile significant
```{r irf}
par(mfrow = c(1,2))
# larceny from vehicle
irf_vehicle1 <- irf(VAR_vehicle,
                    impulse = "larceny_from_vehicle",
                    response = "residuals",
                    n.ahead = 24)
irf_vehicle2 <- irf(VAR_vehicle,
                    impulse = "residuals",
                    response = "larceny_from_vehicle",
                    n.ahead = 24)

# ggplot version of irf
lags = c(1:25)

irf_vehicle_1_gg <- data.frame(irf_vehicle1$irf$larceny_from_vehicle[,1],
                   irf_vehicle1$Lower$larceny_from_vehicle[,1],
                   irf_vehicle1$Upper$larceny_from_vehicle[,1])

colnames(irf_vehicle_1_gg) <- c("mean", "lower", "upper")

ggplot(irf_vehicle_1_gg, aes(x = lags)) +
  geom_line(aes(y = mean), color = "black") +
  #geom_line(aes(y = upper), color = "red", linetype = "dashed") +
  #geom_line(aes(y = lower), color = "red", linetype = "dashed") +
  geom_hline(yintercept = 0, color = "blue") +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = "lightblue", alpha = 0.5) +
  theme_classic() +
  ggtitle("How many more daily covid19 cases there will be after 1 larceny from vehicle") +
  xlab("Number of days after a larceny from vehicle")+
  ylab("Number of covid 19 cases")


irf_vehicle_2_gg <- data.frame(irf_vehicle2$irf$residuals[,1],
                              irf_vehicle2$Lower$residuals[,1],
                              irf_vehicle2$Upper$residuals[,1])

colnames(irf_vehicle_2_gg) <- c("mean", "lower", "upper")

ggplot(irf_vehicle_2_gg, aes(x = lags)) +
  geom_line(aes(y = mean), color = "black") +
  #geom_line(aes(y = upper), color = "red", linetype = "dashed") +
  #geom_line(aes(y = lower), color = "red", linetype = "dashed") +
  geom_hline(yintercept = 0, color = "blue") +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = "lightblue", alpha = 0.5) +
  theme_classic() +
  ggtitle("How many more larceny from vechile cases per day there will be after 1 confirmed covid19 case") +
  xlab("Number of days after a confimed covid 19")+
  ylab("Number of larceny from vechile")
# this one not really significant at a bit above 0.05 and the impact is not even -1.0. So could be okay to leave this out as well.
```

### Step 8: Forecast
```{r var forecast}
forecast <- forecast(VAR_vehicle)

autoplot(forecast$forecast$larceny_from_vehicle)
autoplot(forecast$forecast$residuals)

accu_compare <- data.frame(rbind(accuracy(VAR_vehicle$varresult[[1]])))
rownames(accu_compare) <- c('larceny from vehicle')
kableExtra::kable(accu_compare, format = 'markdown')
```
