---
title: "Seattle"
output: pdf_document
---
```{r library, message = FALSE, warning = FALSE}
library(tidyverse)
library(tseries)
library(tsbox) # transform data into time series
library(xts)
library(COVID19) # to get data about covid 19
library(forecast) #arima model
library(vars) #VAR and Causality
```

# COVID19

## Load COVID 19 data for Chicago
```{r covid 19 for chicago region}
covid19_SEA <- covid19("USA", level = 3) %>%
  # this cook county contains chicago
  filter(administrative_area_level_3 == "King",
         administrative_area_level_2 == "Washington" ) %>%
  # filter out days when confirmed is zero or one
  # becasue when it was 2 for a very long time
  filter(confirmed > 2)

# brief display
head(covid19_SEA)
```
## Overview of the data
### Visualization
```{r overview of covid, message=FALSE}
# plot cumulative cases
# extract for transforming into time series
ts_SEA <- covid19_SEA %>%
  dplyr::select(date, confirmed) %>%
  ts_xts()

plot.xts(ts_SEA,
         main = "Cumulative confirmed cases
         of COVID 19 in Seattle")

# plot daily cases
# first difference
adj_diff_GA <- na.omit(ts_diff_GA[,1] + 7)

plot.xts(ts_diff_SEA,
         main = "Daily confirmed cases of
         COVID19 in Seattle")
```

## Model the infection
#### Construct the model
```{r GAMM model for covid}
# construct GAMM model from prof. Brown's work

# construct data frame of difference, not time series
covid19_SEA_diff <- data.frame(diff(covid19_SEA$confirmed) + 7)
  
colnames(covid19_SEA_diff)[1] = "confirmed"
covid19_SEA_diff$date = covid19_SEA$date[2:length(covid19_SEA$date)]

# time as integer
covid19_SEA_diff$timeInt = as.numeric(covid19_SEA_diff$date)
# RIke a copy to avoid perfect collinearity for mixed effect
covid19_SEA_diff$timeIid = covid19_SEA_diff$timeInt

# GAMM model
gamSEA <- gamm4::gamm4(confirmed ~ s(timeInt, k = 70), 
                      random = ~(1|timeIid),
                      data = covid19_SEA_diff,
                      family = poisson(link = 'log'))
# currently 70 is the max due to length of data
```

#### Obtain the residuals
```{r covid 19 residual}
# obtain forecast
forecast_covid <- data.frame(exp(do.call(cbind, mgcv::predict.gam(gamSEA$gam, toPredict, se.fit=TRUE))))
                        
                        
# access residuals
SEA_res <- data.frame(covid19_GA_diff$confirmed - forecast_covid$fit)

# transform into time series
SEA_res$time = covid19_GA_diff$date
colnames(SEA_res)[1] = "residuals"

col_order <- c("time", "residuals")
SEA_res <- SEA_res[, col_order]

SEA_res_ts <- ts_xts(SEA_res)

plot.xts(SEA_res_ts,
         main = "Daily fluctuation of COVID 19 confirmd cases
         in RI outside its overall trend")
# adf test
adf.test(as.ts(SEA_res_ts))
```

## Load crime data
### 2014 - 2020
```{r get seattle crime data: 2014-2020}
seattle <- read.socrata(
  'https://data.seattle.gov/api/views/tazs-3rd5/rows.csv?accessType=DOWNLOAD',
  app_token = "hPU78MH7zKApdpUv4OVCInPOQ")
seattle <- seattle %>%
  filter(substr(offense_end_datetime, start = 1, stop = 4) >= '2014')

head(seattle)

# many obvious data entry error
seattle$offense_end_datetime[20717] <- '2019-05-28 03:30:00'
seattle$offense_end_datetime[232496] <- '2020-05-08 15:45:00'
seattle$offense_end_datetime[175382] <- '2019-07-22 21:46:00'
seattle$offense_end_datetime[221967] <- '2019-12-28 10:00:00'
seattle$offense_end_datetime[171148] <- '2019-12-23 23:59:00'
seattle$offense_end_datetime[213047] <- '2019-11-29 12:00:00'
seattle$offense_end_datetime[1844] <- '2020-01-13 12:00:00'
seattle$offense_end_datetime[168549] <- '2019-08-14 12:40:00'
seattle$offense_end_datetime[168604] <- '2019-07-31 00:00:00'
seattle$offense_end_datetime[168618] <- '2019-07-28 19:59:00'
seattle$offense_end_datetime[206835] <- '2019-06-07 12:18:00'
seattle$offense_end_datetime[206835] <- '2019-06-07 12:18:00'

# add date
seattle <- seattle %>%
  mutate(y_month  = substr(offense_end_datetime, start = 1, stop = 7)) %>%
  mutate(YEAR  = substr(offense_end_datetime, start = 1, stop = 4)) %>%
  mutate(MONTH = substr(offense_end_datetime, start = 6, stop = 7)) %>%
  mutate(Date = substr(offense_end_datetime, start = 1, stop = 10))
```

```{r}
# summary of all crime
seattle_summary <- seattle %>%
  group_by(offense) %>%
  summarise(number_of_crime = n()) %>%
  arrange(desc(number_of_crime))

# bar chart of 5 most frequent crime over the years
seattle %>%
  filter(offense %in% head(seattle_summary$offense, 5)) %>%
  ggplot(aes(x=offense, fill=offense)) +
  geom_bar(width = 0.5) +
  coord_flip() +
  theme_classic() +
  labs(y='Number of Crimes',x='Offense Description')

# per MONTH
# exclude May
seattle %>%
  dplyr::select(MONTH, offense, YEAR) %>%
  filter(offense %in% seattle_summary$offense[1:5], YEAR == '2020') %>% 
  count(MONTH, offense) %>%
  ggplot(aes(MONTH, n, group = offense, color = offense)) +
  geom_line() +
  scale_color_brewer(palette = "Set1") +
  ggtitle("Monthly frequency of top 5 crime in Seattle in 2020")

# This dataset doesn't have 'day of the week' yet.

# daily
seattle %>%
  dplyr::select(offense_end_datetime, offense, YEAR) %>%
  filter(offense %in% seattle_summary$offense[1:5], YEAR=='2020') %>%
  count(offense_end_datetime, offense) %>%
  ggplot(aes(offense_end_datetime, n, group = offense, color = offense)) +
  geom_line() +
  facet_wrap(~offense) +
  scale_fill_brewer(palette = "Set1", breaks = rev(levels(seattle_summary$offense[1:5]))) +
  ggtitle("Frequency of top 5 crime in Seattle in 2020")

# year to year comparison
seattle %>%
  dplyr::select(y_month, MONTH, offense, YEAR) %>%
  filter(offense %in% seattle_summary$offense[1:5]) %>%
  count(YEAR, MONTH, offense) %>%
  na.omit() %>%
  ggplot(aes(x=MONTH, y=n, group = YEAR, color = as.character(YEAR))) +
  geom_line() +
  facet_wrap(~offense) +
  ggtitle("year to year comparison of the top 5 crime in Seattle in the past 6 years")
```

## VAR 
### Step1 : Extract cases
```{r extract cases}
# extract top 5 crime
top5crime <- seattle %>%
  filter(offense %in% head(seattle_summary$offense, 5)) %>%
  group_by(Date, offense) %>%
  tally() %>%
  spread(offense, n)

# rename columns
colnames(top5crime) <- c('time',
                         "larceny",
                         "burglary",
                         "vandalism",
                         'vehicle_theft',
                         "theft_from_vehicle")
top5crime <- na.omit(top5crime)
# create time series
top5crime_xts <- ts_xts(top5crime[,1:2])

for (i in (3:ncol(top5crime))){
  temp_xts <- ts_xts(top5crime[, c(1,i)])
  top5crime_xts <- merge(top5crime_xts, temp_xts)
}

# extract difference, change per day
top5crime_diff <- na.omit(diff(top5crime_xts))
```
### Step 2: Construct combined time series
```{r top 5 crime VAR}
# specify common time range
# start from when covid was a thing
# end with 1 day before today's date
common_time <- seq.Date(start(SEA_res_ts), Sys.Date() - 1 , by = "day")

# combine time series of crime and covid
combined_diff <- merge(top5crime_diff[paste(common_time[1],
                                            common_time[length(common_time)],
                                            sep = "/")],
                       SEA_res_ts[paste(common_time[1],
                                            common_time[length(common_time)],
                                            sep = "/")])

```

### Step 3: Plot each crime with covid
```{r plot together}
for (i in 1:(ncol(combined_diff) - 1)){
  plotrix::twoord.plot(common_time,
                       combined_diff[,i],
                       common_time,
                       combined_diff$residuals,
                       type = c("l","l"),
                       xaxt = "n",
                       rylab = "number of daily fluctuation of covid 19 cases",
                       ylab = paste("daily change in", colnames(combined_diff)[i]))
                       
}
```

### Step 5: Construct VAR model
```{r construct var, warning = FALSE}
# variable selection based on AIC
optimal_larceny <- VARselect(na.omit(combined_diff)[,c(1,6)], type = 'none', lag.max = 10)
optimal_burglary <- VARselect(na.omit(combined_diff)[,c(2,6)], type = 'none', lag.max = 10)
optimal_vandalism <- VARselect(na.omit(combined_diff)[,c(3,6)], type = 'none', lag.max = 10)
optimal_vehicle_theft <- VARselect(na.omit(combined_diff)[,c(4,6)], type = 'none', lag.max = 10)
optimal_theft_fromvehicle <- VARselect(na.omit(combined_diff)[,c(5,6)], type = 'none', lag.max = 10)

# use AIC as selection criteria
VAR_larceny <- VAR(y=as.ts(na.omit(combined_diff)[,c(1,6)]), p=optimal_assault$selection[1])
VAR_burglary <- VAR(y=as.ts(na.omit(combined_diff)[,c(2,6)]),
                     p=optimal_autotheft$selection[1])
VAR_vandalism <- VAR(y=as.ts(na.omit(combined_diff)[,c(3,6)]),
                    p=optimal_burglary$selection[1])
VAR_vehicle_theft <- VAR(y=as.ts(na.omit(combined_diff)[,c(4,6)]),
                               p=optimal_larceny_fromvehicle$selection[1])
VAR_theft_fromvehicle<- VAR(y=as.ts(na.omit(combined_diff)[,c(5,6)]),
                              p=optimal_larceny_nonvehicle$selection[1])
```

### Step 6: Granger Causality test

```{r}
causality(VAR_larceny, cause = colnames(combined_diff)[1])
causality(VAR_larceny, cause = "residuals")
```

```{r}
causality(VAR_burglary, cause = colnames(combined_diff)[2])
causality(VAR_burglary, cause = "residuals")
```

burglary significant to covid-19 (p-value = 0.0425)
instantaneous causality between: burglary and residuals (p-value = 0.001079)

```{r}
causality(VAR_vandalism, cause = colnames(combined_diff)[3])
causality(VAR_vandalism, cause = "residuals")
```

covid-19 significant to vandalism (p-value = 0.004843)

```{r}
causality(VAR_vehicle_theft, cause = colnames(combined_diff)[4])
causality(VAR_vehicle_theft, cause = "residuals")
```

```{r}
causality(VAR_theft_fromvehicle, cause = colnames(combined_diff)[5])
causality(VAR_theft_fromvehicle, cause = "residuals")
```
weak instantaneous causality between: theft_from_vehicle and residuals (p-value = 0.06536)

### Step 7: Impulse Response Function

Only vandalism significant
```{r irf}
par(mfrow = c(1,2))
# larceny from vehicle
irf_vandalism1 <- irf(VAR_vandalism,
                    impulse = 'vandalism',
                    response = "residuals",
                    n.ahead = 24)
irf_vandalism2 <- irf(VAR_vandalism,
                    impulse = "residuals",
                    response = "vandalism",
                    n.ahead = 24)

plot(irf_vandalism1)
plot(irf_vandalism2)

ir.c1 <- irf(VAR_vandalism, 
                         impulse = "vandalism", 
                         response = "residuals", 
                         n.ahead = 24,
                         ortho = F,
                         cumulative = T)
plot(ir.c1)

ir.c2 <- irf(VAR_vandalism, 
                         impulse = "residuals", 
                         response = "vandalism", 
                         n.ahead = 24,
                         ortho = F,
                         cumulative = T)
plot(ir.c2)
```
### Step 8: Forecast
```{r var forecast}
forecast(VAR_vandalism) %>%
  autoplot()


accu_compare <- data.frame(rbind(accuracy(VAR_vandalism$varresult[[1]])))
rownames(accu_compare) <- c('vandalism')
kableExtra::kable(accu_compare, format = 'markdown')
```