---
title: "Seattle"
output: pdf_document
---
```{r library, message = FALSE, warning = FALSE}
library(RSocrata)
library(tidyverse)
library(tseries)
library(tsbox) # transform data into time series
library(xts)
library(COVID19) # to get data about covid 19
library(forecast) #arima model
library(vars) #VAR and Causality
library(plotly)
library(dygraphs)
```

# COVID19

## Load COVID 19 data for Chicago
```{r covid 19 for chicago region}
covid19_SEA <- covid19("USA", level = 3) %>%
  # this cook county contains chicago
  filter(administrative_area_level_3 == "King",
         administrative_area_level_2 == "Washington" ) %>%
  # filter out days when confirmed is zero or one
  # becasue when it was 1 for a very long time
  filter(confirmed > 1)

# brief display
head(covid19_SEA)
```
## Overview of the data
### Visualization
```{r overview of covid, message=FALSE}
# plot cumulative cases
# extract for transforming into time series
ts_SEA <- covid19_SEA %>%
  dplyr::select(date, confirmed) %>%
  ts_xts()

plot.xts(ts_SEA,
         main = "Cumulative confirmed cases
         of COVID 19 in Seattle")

# plot daily cases
# first difference
ts_diff_SEA <- na.omit(diff(ts_SEA))

plot.xts(ts_diff_SEA,
         main = "Daily confirmed cases of
         COVID19 in Seattle")
```

## Model the infection
#### Construct the model
```{r GAMM model for covid}
# construct GAMM model from prof. Brown's work

# construct data frame of difference, not time series
# one day has -7 daily confirmed cases
# add back 7 to all
# does not affect residuals after GAMM
covid19_SEA_diff <- data.frame(diff(covid19_SEA$confirmed) - min(ts_diff_SEA))
  
colnames(covid19_SEA_diff)[1] = "confirmed"
covid19_SEA_diff$date = covid19_SEA$date[2:length(covid19_SEA$date)]

# time as integer
covid19_SEA_diff$timeInt = as.numeric(covid19_SEA_diff$date)
# RIke a copy to avoid perfect collinearity for mixed effect
covid19_SEA_diff$timeIid = covid19_SEA_diff$timeInt

# GAMM model
gamSEA <- gamm4::gamm4(confirmed ~ s(timeInt, k = 90), 
                      random = ~(1|timeIid),
                      data = covid19_SEA_diff,
                      family = poisson(link = 'log'))
```

#### Obtain the residuals
```{r covid 19 plot GAMM}
# plot fitted value
toPredict = data.frame(time = seq(covid19_SEA_diff$date[1],
                                  covid19_SEA_diff$date[length(covid19_SEA_diff$date)],
                                  by = '1 day'))

toPredict$timeInt = as.numeric(toPredict$time)

matplot(toPredict$time, 
        exp(do.call(cbind, mgcv::predict.gam(gamSEA$gam, toPredict, se.fit=TRUE)) %*% 
              Pmisc::ciMat()), 
        col='red', lty=c(1,2,2), type='l', xaxt='n', xlab='', ylab='Daliy Confirmed cases', 
        ylim = c(0.5, 500), xlim = as.Date(c(covid19_SEA$date[1], covid19_SEA$date[length(covid19_SEA$date)])))
title("Daily confirmed cases of COVID 19 in Seattle")
matpoints(toPredict$time, covid19_SEA_diff$confirmed, 
          col = 'black',
          type = 'l')

```


```{r covid 19 residuals}
# obtain forecast
forecast_covid <- data.frame(exp(do.call(cbind, mgcv::predict.gam(gamSEA$gam, toPredict, se.fit=TRUE))))
                        
                        
# access residuals
SEA_res <- data.frame(covid19_SEA_diff$confirmed - forecast_covid$fit)

# transform into time series
SEA_res$time = covid19_SEA_diff$date
colnames(SEA_res)[1] = "residuals"

col_order <- c("time", "residuals")
SEA_res <- SEA_res[, col_order]

SEA_res_ts <- ts_xts(SEA_res)

plot.xts(SEA_res_ts,
         main = "Daily fluctuation of COVID 19 confirmd cases
         in RI outside its overall trend")
# adf test
adf.test(as.ts(SEA_res_ts))
```

## Load crime data
### 2014 - 2020
```{r get seattle crime data: 2014-2020}
seattle <- read.socrata(
  'https://data.seattle.gov/api/views/tazs-3rd5/rows.csv?accessType=DOWNLOAD',
  app_token = "hPU78MH7zKApdpUv4OVCInPOQ")

seattle <- seattle %>%
  filter(substr(report_datetime, start = 1, stop = 4) >= '2014')

head(seattle)

# add date
seattle <- seattle %>%
  mutate(y_month  = substr(report_datetime, start = 1, stop = 7)) %>%
  mutate(YEAR  = substr(report_datetime, start = 1, stop = 4)) %>%
  mutate(MONTH = substr(report_datetime, start = 6, stop = 7)) %>%
  mutate(Date = substr(report_datetime, start = 1, stop = 10))
```

```{r}
# summary of all crime
seattle_summary <- seattle %>%
  group_by(offense) %>%
  summarise(number_of_crime = n()) %>%
  arrange(desc(number_of_crime))

# bar chart of 5 most frequent crime over the years
seattle %>%
  filter(offense %in% head(seattle_summary$offense, 5)) %>%
  ggplot(aes(x=offense, fill=offense)) +
  geom_bar(width = 0.5) +
  coord_flip() +
  theme_classic() +
  labs(y='Number of Crimes',x='Offense Description')

# per MONTH in 2020
# exclude June
seattle %>%
  dplyr::select(MONTH, offense, YEAR) %>%
  filter(offense %in% seattle_summary$offense[1:5], YEAR == '2020', MONTH != "06") %>% 
  count(MONTH, offense) %>%
  ggplot(aes(MONTH, n, group = offense, color = offense)) +
  geom_line() +
  scale_color_brewer(palette = "Set1") +
  ggtitle("Monthly frequency of top 5 crime in Seattle in 2020")

# This dataset doesn't have 'day of the week' yet.

# daily
daily <- seattle %>%
  dplyr::select(Date, offense, YEAR) %>%
  filter(offense %in% seattle_summary$offense[1:5], YEAR=='2020') %>%
  count(Date, offense) %>%
  ggplot(aes(Date, n, group = offense, color = offense)) +
  geom_line() +
  facet_free(~offense) +
  scale_fill_brewer(palette = "Set1", breaks = rev(levels(seattle_summary$offense[1:5]))) +
  ggtitle("Frequency of top 5 crime in Seattle in 2020") +
  theme(legend.position = "none")

ggplotly(daily)

# year to year comparison
# exclude 2020-06
yty <- seattle %>%
  dplyr::select(y_month, MONTH, offense, YEAR) %>%
  filter(offense %in% seattle_summary$offense[1:5],
         y_month != "2020-06") %>%
  count(YEAR, MONTH, offense) %>%
  na.omit() %>%
  ggplot(aes(x=MONTH, y=n, group = YEAR, color = as.character(YEAR))) +
  geom_line() +
  facet_free(~offense) +
  guides(color = guide_legend(reverse = TRUE)) +
  ggtitle("year to year comparison of the top 5 crime in Seattle in the past 6 years") +
  labs(col='Year')

ggplotly(yty) %>%
  layout(legend=list(traceorder='reversed'))
```

## VAR 
### Step1 : Extract cases
```{r extract cases}
# extract top 5 crime
top5crime <- seattle %>%
  filter(offense %in% head(seattle_summary$offense, 5)) %>%
  group_by(Date, offense) %>%
  tally() %>%
  spread(offense, n)

# rename columns
colnames(top5crime) <- c('time',
                         "larceny",
                         "burglary",
                         "vandalism",
                         'vehicle_theft',
                         "theft_from_vehicle")
top5crime <- na.omit(top5crime)
# create time series
top5crime_xts <- ts_xts(top5crime[,1:2])

for (i in (3:ncol(top5crime))){
  temp_xts <- ts_xts(top5crime[, c(1,i)])
  top5crime_xts <- merge(top5crime_xts, temp_xts)
}

# extract difference, change per day
top5crime_diff <- na.omit(diff(top5crime_xts))
```
### Step 2: Construct combined time series
```{r top 5 crime VAR}
# specify common time range
# start from when covid was a thing
# end with 1 day before today's date
common_time <- seq.Date(start(SEA_res_ts), as.Date("2020-05-25") , by = "day")

# combine time series of crime and covid
combined_diff <- merge(top5crime_diff[paste(common_time[1],
                                            common_time[length(common_time)],
                                            sep = "/")],
                       SEA_res_ts[paste(common_time[1],
                                            common_time[length(common_time)],
                                            sep = "/")])

```

### Step 3: Plot each crime with covid
```{r plot together}
for (i in 1:(ncol(combined_diff) - 1)){
  plotrix::twoord.plot(common_time,
                       combined_diff[,i],
                       common_time,
                       combined_diff$residuals,
                       type = c("l","l"),
                       xaxt = "n",
                       rylab = "number of daily fluctuation of covid 19 cases",
                       ylab = paste("daily change in", colnames(combined_diff)[i]))
                       
}
```

### Step 5: Construct VAR model
```{r construct var, warning = FALSE}
# variable selection based on AIC
optimal_larceny <- VARselect(na.omit(combined_diff)[,c(1,6)], type = 'none', lag.max = 10)
optimal_burglary <- VARselect(na.omit(combined_diff)[,c(2,6)], type = 'none', lag.max = 10)
optimal_vandalism <- VARselect(na.omit(combined_diff)[,c(3,6)], type = 'none', lag.max = 10)
optimal_vehicle_theft <- VARselect(na.omit(combined_diff)[,c(4,6)], type = 'none', lag.max = 10)
optimal_theft_fromvehicle <- VARselect(na.omit(combined_diff)[,c(5,6)], type = 'none', lag.max = 10)

# use AIC as selection criteria
VAR_larceny <- VAR(y=as.ts(na.omit(combined_diff)[,c(1,6)]),
                   p=optimal_larceny$selection[1])
VAR_burglary <- VAR(y=as.ts(na.omit(combined_diff)[,c(2,6)]),
                     p=optimal_burglary$selection[1])
VAR_vandalism <- VAR(y=as.ts(na.omit(combined_diff)[,c(3,6)]),
                    p=optimal_vandalism$selection[1])
VAR_vehicle_theft <- VAR(y=as.ts(na.omit(combined_diff)[,c(4,6)]),
                         p=optimal_vehicle_theft$selection[1])
VAR_theft_fromvehicle<- VAR(y=as.ts(na.omit(combined_diff)[,c(5,6)]),
                              p=optimal_theft_fromvehicle$selection[1])
```

### Step 6: Granger Causality test
#### LARCENY
```{r}
causality(VAR_larceny, cause = colnames(combined_diff)[1])
causality(VAR_larceny, cause = "residuals")
```
Nothing significant

#### BURGLARY
```{r}
causality(VAR_burglary, cause = colnames(combined_diff)[2])
causality(VAR_burglary, cause = "residuals")
```

covid significant to burglary (0.008)

#### VANDALSIM
```{r}
causality(VAR_vandalism, cause = colnames(combined_diff)[3])
causality(VAR_vandalism, cause = "residuals")
```

Nothing significant

#### VEHICLE THEFT
```{r}
causality(VAR_vehicle_theft, cause = colnames(combined_diff)[4])
causality(VAR_vehicle_theft, cause = "residuals")
```

Nothing significant.

#### THEFT FROM VEHICLE
```{r}
causality(VAR_theft_fromvehicle, cause = colnames(combined_diff)[5])
causality(VAR_theft_fromvehicle, cause = "residuals")
```
Nothing significant.

### Step 7: Impulse Response Function

Only vandalism significant
```{r irf}
lags = c(1:25)

par(mfrow = c(1,2))
# only covid significant to bulglary
irf_burglary_1 <- irf(VAR_burglary,
                      impulse = "residuals",
                      response = "burglary",
                      n.ahead = 24)
# ggplot
irf_burglary_1_gg <- data.frame(irf_burglary_1$irf$residuals[,1],
                                irf_burglary_1$Lower$residuals[,1],
                                irf_burglary_1$Upper$residuals[,1])

colnames(irf_burglary_1_gg) <- c("mean", "lower", "upper")

irf_burglary_1_plot <- ggplot(irf_burglary_1_gg, aes(x=lags)) +
  geom_line(aes(y = mean), color = "black") +
  geom_hline(yintercept = 0, color = "blue") +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = "lightblue", alpha = 0.5) +
  theme_classic() +
  ggtitle("How many more burglary cases per day there will be 
          after 1 confirmed covid19 case") +
  xlab("Number of days after a confimed covid 19 case")+
  ylab("Number of bulglary cases")

ggplotly(irf_burglary_1_plot)
```
### Step 8: Forecast
```{r custom function}
interval_value_formatter <- "function(num, opts, seriesName, g, row, col) {
  value = g.getValue(row, col);
  if(value[0] != value[2]) {
    lower = Dygraph.numberValueFormatter(value[0], opts);
    upper = Dygraph.numberValueFormatter(value[2], opts);
    return '[' + lower + ', ' + upper + ']';
  } else {
    return Dygraph.numberValueFormatter(num, opts);
  }
}"
```

```{r var forecast}
# larceny
# not significant
forecast_larceny <- forecast(VAR_larceny)

forecast_larceny$forecast$larceny %>%
  {cbind(actuals = .$x, forecast_mean=.$mean,
         lower_95=.$lower[,"95%"], upper_95=.$upper[,"95%"],
         lower_80=.$lower[,"80%"], upper_80=.$upper[,"80%"])} %>%
  dygraph(main = "Daily forecast for day-to-day change in larceny in Seattle",
          ylab = "Day-to-day change") %>%
  dyAxis("y", valueFormatter = interval_value_formatter) %>%
  dySeries("actuals", color = "black") %>%
  dySeries("forecast_mean", color = "blue", label = "forecast") %>%
  dySeries(c("lower_80", "forecast_mean", "upper_80"),
           label = "80%", color = "blue") %>%
  dySeries(c("lower_95", "forecast_mean", "upper_95"),
           label = "95%", color = "blue")%>%
  dyLegend(labelsSeparateLines=TRUE) %>%
  dyRangeSelector() %>%
  dyAxis("x", label = paste("Numebr of days since", common_time[1])) %>%
  dyLegend(show = "follow")

# burglary
# covid significant to crime
forecast_burglary <- forecast(VAR_burglary)

forecast_burglary$forecast$burglary %>%
  {cbind(actuals = .$x, forecast_mean=.$mean,
         lower_95=.$lower[,"95%"], upper_95=.$upper[,"95%"],
         lower_80=.$lower[,"80%"], upper_80=.$upper[,"80%"])} %>%
  dygraph(main = "Daily forecast for day-to-day change in burglary in Seattle",
          ylab = "Day-to-day change") %>%
  dyAxis("y", valueFormatter = interval_value_formatter) %>%
  dySeries("actuals", color = "black") %>%
  dySeries("forecast_mean", color = "blue", label = "forecast") %>%
  dySeries(c("lower_80", "forecast_mean", "upper_80"),
           label = "80%", color = "blue") %>%
  dySeries(c("lower_95", "forecast_mean", "upper_95"),
           label = "95%", color = "blue")%>%
  dyLegend(labelsSeparateLines=TRUE) %>%
  dyRangeSelector() %>%
  dyAxis("x", label = paste("Numebr of days since", common_time[1])) %>%
  dyLegend(show = "follow")

# vandalism
# not significant
forecast_vandalism <- forecast(VAR_vandalism)

forecast_vandalism$forecast$vandalism %>%
  {cbind(actuals = .$x, forecast_mean=.$mean,
         lower_95=.$lower[,"95%"], upper_95=.$upper[,"95%"],
         lower_80=.$lower[,"80%"], upper_80=.$upper[,"80%"])} %>%
  dygraph(main = "Daily forecast for day-to-day change in vandalism in Seattle",
          ylab = "Day-to-day change") %>%
  dyAxis("y", valueFormatter = interval_value_formatter) %>%
  dySeries("actuals", color = "black") %>%
  dySeries("forecast_mean", color = "blue", label = "forecast") %>%
  dySeries(c("lower_80", "forecast_mean", "upper_80"),
           label = "80%", color = "blue") %>%
  dySeries(c("lower_95", "forecast_mean", "upper_95"),
           label = "95%", color = "blue")%>%
  dyLegend(labelsSeparateLines=TRUE) %>%
  dyRangeSelector() %>%
  dyAxis("x", label = paste("Numebr of days since", common_time[1])) %>%
  dyLegend(show = "follow")

# vehicle theft
# not significant
forecast_vehicle_theft <- forecast(VAR_vehicle_theft)

forecast_vehicle_theft$forecast$vehicle_theft %>%
  {cbind(actuals = .$x, forecast_mean=.$mean,
         lower_95=.$lower[,"95%"], upper_95=.$upper[,"95%"],
         lower_80=.$lower[,"80%"], upper_80=.$upper[,"80%"])} %>%
  dygraph(main = "Daily forecast for day-to-day change in vehicle theft in Seattle",
          ylab = "Day-to-day change") %>%
  dyAxis("y", valueFormatter = interval_value_formatter) %>%
  dySeries("actuals", color = "black") %>%
  dySeries("forecast_mean", color = "blue", label = "forecast") %>%
  dySeries(c("lower_80", "forecast_mean", "upper_80"),
           label = "80%", color = "blue") %>%
  dySeries(c("lower_95", "forecast_mean", "upper_95"),
           label = "95%", color = "blue")%>%
  dyLegend(labelsSeparateLines=TRUE) %>%
  dyRangeSelector() %>%
  dyAxis("x", label = paste("Numebr of days since", common_time[1])) %>%
  dyLegend(show = "follow")

# theft from vehicle
# not significant
forecast_theft_fromvehicle <- forecast(VAR_theft_fromvehicle)

forecast_theft_fromvehicle$forecast$theft_from_vehicle %>%
  {cbind(actuals = .$x, forecast_mean=.$mean,
         lower_95=.$lower[,"95%"], upper_95=.$upper[,"95%"],
         lower_80=.$lower[,"80%"], upper_80=.$upper[,"80%"])} %>%
  dygraph(main = "Daily forecast for day-to-day change in theft from vehicle in Seattle",
          ylab = "Day-to-day change") %>%
  dyAxis("y", valueFormatter = interval_value_formatter) %>%
  dySeries("actuals", color = "black") %>%
  dySeries("forecast_mean", color = "blue", label = "forecast") %>%
  dySeries(c("lower_80", "forecast_mean", "upper_80"),
           label = "80%", color = "blue") %>%
  dySeries(c("lower_95", "forecast_mean", "upper_95"),
           label = "95%", color = "blue")%>%
  dyLegend(labelsSeparateLines=TRUE) %>%
  dyRangeSelector() %>%
  dyAxis("x", label = paste("Numebr of days since", common_time[1])) %>%
  dyLegend(show = "follow")


accu_compare <- data.frame(rbind(accuracy(VAR_vandalism$varresult[[1]])))
rownames(accu_compare) <- c('vandalism')
kableExtra::kable(accu_compare, format = 'markdown')
```
